---
title: "ETO5513 Assignment 1"
output: html_document
date: "2023-05-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
options(readr.show_col_types = FALSE)
library(knitr)
library(dplyr)
#library(tidyverse)
#library(visdat)
#library(ggplot2)
#library(gganimate)
#library(plotly)
#library(gapminder)
```

## Research Question
The data being examined in this assignment is to investigate the relationship between Australia's median age and birth rate trends. This is to determine whether the decrease in birth rates have contributed to Australia's aging population.

## Dataset Introduction
The dataset "population-growth-rate-vs-median-age.csv" is sourced from https://ourworldindata.org/grapher/population-growth-rate-vs-median-age. The dataset contains 5 columns:
Entity: The country
Code: Country code relating to Entity
Year: The year relating to the data point
NATURAL GROWTH RATE - SEX: ALL - AGE: ALL - VARIANT: ESTIMATES: All - Age: All - Variant: Estimates: Crude birth rate minus the crude death rate. Represents the portion of population growth (or decline) determined exclusively by births and deaths.
MEDIAN AGE - SEX: ALL - AGE: ALL - VARIANT: ESTIMATES: Age that divides the population in two parts of equal size, that is, there are as many persons with ages above the median as there are with ages below the median.
Continent: The continent relating to the data point.

```{r, include=FALSE}
tb <- read_csv("population-growth-rate-vs-median-age.csv")
tb <- filter(tb, Entity=="Australia")
```

```{r}
kable(tb)
```